---
title: "Data Analysis of Tech. Salaries at Amazon and Google"
author: "The Statisticians"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: readable
    toc: yes
urlcolor: cyan
---

### <u>Introduction</u>

As data science students, we're intrigued by the compensation structures within leading tech companies like Amazon and Google. To delve into this interest, we're undertaking a project to analyze a dataset of 12,647 salary records from these companies, with an aim to construct a predictive model for annual total compensation.

This dataset, `Levels-Fyi-Salary-Data-Cleaned.csv`, provides a unique and in-depth perspective into the salary structures of numerous top-tier companies. The dataset encompasses 28 variables, covering a variety of aspects such as company, level, title, location, years of experience, years at the company, and various demographic factors including gender and education level. Of particular interest to us is the `totalyearlycompensation` variable, which will serve as the response variable in our analysis.

Our interest in this dataset is driven by our curiosity about the factors that influence annual compensation in top companies. We aim to identify the key variables and their interactions that significantly impact salary levels. This would provide insights that could be invaluable to a variety of stakeholders, from job seekers and employees to human resource professionals and policymakers.

The primary goal of creating a model with this data is to predict the `totalyearlycompensation` based on a set of predictor variables. This model will help us understand how different variables, including categorical ones like `company` and numeric ones such as `yearsofexperience` and `basesalary`, contribute to the total yearly compensation of an employee. Through this process, we also hope to identify potential trends, anomalies, or points of interest that could provide further avenues for exploration and analysis.

***

### <u>Methods</u>

-- We choose the following model... `(PLACE HOLDER)`

```{r message=FALSE, warning=FALSE}
# Loading dataset
library(readr)
data = read_csv("Levels-Fyi-Salary-Data-Cleaned.csv")

# Subsetting dataset and changing some predictors to factors

```

```{r}
# Placeholder and we will just put the model we actually use here 
model = 2
```


***

### <u>Results</u>


***

### <u>Discussion</u>


***

### <u>Appendix</u> 

Model 1
```{r}

```

Model 2 
```{r}

```


Model 3 
```{r}

```


By: Nicholas O. Brown, Ellie Jung, Angela Mei

***

